\documentclass[11pt]{manual}
\usepackage[utf8]{inputenc}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
%\usepackage{amssymb}
\usepackage{makeidx}
\usepackage{url}
\usepackage{colortbl}
\usepackage[table]{xcolor}
\definecolor{seagreen}{RGB}{87,135,159}
\definecolor{light-gray}{gray}{0.85}
\definecolor{medium-gray}{gray}{0.5}
\usepackage[colorlinks=true,linkcolor= seagreen]{hyperref}

\title{The Trampoline handbook}
\author{Jean-Luc B\'echennec}

\release{2.0}
\makeindex

\begin{document}
\maketitle
~\newpage
\setcounter{tocdepth}{2}
\tableofcontents

\chapter{Compiling Trampoline}

\section{Available flags}

Trampoline can be compiled with various options. These options are controlled by setting the appropriate flags. The following flags are availables:
\begin{description}
\item[OS\_EXTENDED] is used to compile Trampoline with extended error support according to the specification.
\item[WITH\_ERROR\_HOOK] adds the call to the error hook when an error occurs in an os system call.
\item[WITH\_PRE\_TASK\_HOOK] adds the call to the pre-task hook function.
\item[WITH\_POST\_TASH\_HOOK]  adds the call to the post-task hook function.
\item[WITH\_TASK\_HOOK] is the same as selecting the two previous flags.
\item[WITH\_STARTUP\_HOOK] adds the call to the startup hook function.
\item[WITH\_SHUTDOWN\_HOOK] adds the call to the shutdown hook function.
\item[WITH\_INTERNAL\_EXTERNAL\_COM] is used when a same sending message object does an internal \emph{and} an external communication.
\item[WITH\_COM\_ERROR\_HOOK] adds the call to the communication error hook when an error occurs in a communication system call.
\end{description}

\input{tasks}
\input{ports}
\input{tp}

 \end{document} 